Improving the Replicability of Psychological Science Through Pedagogy

### 2017-05-17 paper/paper.Rmd

Our projects had a number of limitations, including at least: 1) limited time for iterative piloting and adjustment, 2) limits on funding for larger samples, and 3) limits on domain expertise with respect to the specific effects that were chosen. But within these limitations, our group was able to produce a set of relatively close replications with generally good statistical power. Thus, our work here shows what is possible for motivated students using freely-available tools and resources. Such classes could become the backbone of future collaborative replication efforts [@frank2012;@everett2015].

One of the barriers to more reproducible work in psychology is pedagogical: Some of the tools (programmatic statistical analysis, version control) can feel inaccessible or intimidating to beginners. Indeed, our students, who came from a variety of disciplinary backgrounds, had relatively little experience with web-based empirical studies or literate, reproducible data analysis. Yet by the end of the course, all had gained sufficient proficiency with the suite of technical and conceptual tools necessary to carry out a study independently. This experience suggests that it is possible to convey key concepts in sufficient depth that -- within a supportive environment -- students can quickly learn to use a reproducible toolset.

From a purely statistical point of view, though, our results were underwhelming. Despite our relatively close adherence to the original protocols and relatively large sample sizes, the modal outcome of our projects was partial replication. These partial replications often were cases in which some hint of the original pattern was observed but the key statistical test was not statistically significant and showed both smaller effect size and lower evidential value than the same test in the original. We invite readers to browse the narrative descriptions of individual replication attempts in our SOM to see the ways that patterns of empirical findings can differ from one another beyond the significance of a single test.

>>>>

## Pedagogical Implications

Because they were conducted by students within the constraints of a course, our studies had a number of limitations, including at least: 1) limited time for iterative piloting and adjustment, 2) limits on funding for larger samples, and 3) limits on domain expertise with respect to the specific effects that were chosen. But within these limitations, our group was able to produce a set of relatively close replications with generally good statistical power. Thus, our work here shows what is possible for motivated students using freely-available tools and resources. Such classes could become the backbone of future collaborative replication efforts [@frank2012;@everett2015].

While every course of this type will likely be tailored to the interests and instructional needs of a particular department and institution, there will also likely be some commonalities. To encourage others to share our materials, our course outline, project templates, and assignments are available publicly at [https://osf.io/98ta4](https://osf.io/98ta4). In addition, all of the most recent lecture slides and materials are avaialable on our course website at [http://psych254.stanford.edu](). We hope that others will share and reuse our materials as the consider how best to design a course customized to their context.

One of the barriers to more reproducible work in psychology is technical: Some of the tools (programmatic statistical analysis, version control) can feel inaccessible or intimidating to beginners. Indeed, our students, who came from a variety of disciplinary backgrounds, had relatively little experience with web-based empirical studies or literate, reproducible data analysis. Yet by the end of the course, all had gained sufficient proficiency with the suite of technical and conceptual tools necessary to carry out a study independently. This experience suggests that it is possible to convey key concepts in sufficient depth that -- within a supportive environment -- students can quickly learn to use a reproducible toolset.

We end with a few recommendations for instructors designing such a course. First, plan as much as possible to make the project the centerpiece of the course, so that students begin the process as quickly as possible and do the maximal amount of skill learning in the context of their own goal-directed research. Second, rather than adopting the typical model of students turning in work at the end of the course and being assessed, consider a more collaborative model where the instructors provide iterative feedback on subcomponents of projects (e.g., sample analyses, parts of the experimental paradigm or materials). Such feedback can both save students time and frustration and provide a check on project quality before data are collected. Third and inally, try to share the excitement of discovery that is inherent in doing new research! We have found that the most rewarding part of teaching this class is the fun of getting to see the results of all the hard work at the end.

## Implications for Replication Research

From a purely statistical point of view, the results of our studies were underwhelming. Despite our relatively close adherence to the original protocols and relatively large sample sizes, the modal outcome of our projects was partial replication. These partial replications often were cases in which some hint of the original pattern was observed but the key statistical test was not statistically significant and showed both smaller effect size and lower evidential value than the same test in the original. We invite readers to browse the narrative descriptions of individual replication attempts in our SOM to see the ways that patterns of empirical findings can differ from one another beyond the significance of a single test.
